---
title: "Exploring Gun Deaths in America"
output: github_document
---

## Objectives
* To explore gun death data curated by FiveThirtyEight [original data available here](https://github.com/fivethirtyeight/guns-data) by
    * Month
    * Intent
    * Age and sex
    * Education
    * Season
    * Intent by race
    * Police-relatedness 

## Read in the data

The gun_deaths.csv is provided in the repository. 
```{r load-data, message=FALSE, warning=FALSE}

# read in the data

```

## Generating a data frame that summarizes the number of gun deaths per month and printing using `kable()`. 

Knitr is part of the knitr package. Allows you to change column names and add captions and make pretty tables in your knitted document. Hint: set the argument format = "markdown"



### Generating a bar chart with human-readable labels on the x-axis. That is, each month should be labeled "Jan", "Feb", "Mar" (full or abbreviated month names are fine), not `1`, `2`, `3`.

```{r month_chart}
library(tidyverse)  
library(ggplot2) # load tidyverse packages, including ggplot2
library(readr)
library(knitr)
gun_deaths <- read_csv("gun_deaths.csv")
data("gun_deaths")
ggplot(gun_deaths, aes(x = month)) +
 geom_bar(aes(fill = "black"), position = position_dodge()) +
 labs(title="Number of Gun Deaths per Month", 
           x="Month",
       y="Count") +
            scale_x_discrete(labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) 

    
```

## Generating a bar chart that identifies the number of gun deaths associated with each type of intent cause of death. The bars should be sorted from highest to lowest values.

```{r intent}
library(tidyverse)  
library(ggplot2) # load tidyverse packages, including ggplot2
library(readr)
library(knitr)
gun_deaths <- read_csv("gun_deaths.csv")
data("gun_deaths")

gun_deaths %>%
    arrange(intent) %>%
ggplot(gun_deaths, mapping = aes(x = intent)) +
 geom_bar(aes(fill = "black")) +
 labs(title="Number of Gun Deaths by Intent", 
           x="Intent",
       y="Count") 
```

## Generating a boxplot visualizing the age of gun death victims, by sex. Print the average age of female gun death victims.

```{r age_&_sex}
library(tidyverse)    # load tidyverse packages, including ggplot2
library(readr)
library(knitr)
gun_deaths <- read_csv("gun_deaths.csv") 
data("gun_deaths") 
ggplot(gun_deaths, aes(x = sex)) +
 geom_boxplot(aes(fill = "black"))  +
 labs(title="Number Gun Deaths by Sex", 
           x="Sex",
       y="Count")
```


## How many white males with at least a high school education were killed by guns in 2012?

```{r education}
library(ggplot2) # load tidyverse packages, including ggplot2
library(readr)
library(knitr)
gun_deaths <- read_csv("gun_deaths.csv") 
data("gun_deaths") 
gun_deaths %>%
    select(year, race, education)
    group_by(education)
ggplot(gun_deaths, aes(x = sex, y = age)) +
 geom_boxplot(aes(fill = "black"))  +
 theme(legend.position =none) +
 labs(title="Number Gun Death Victims by Intent", 
           x="Intent",
       y="Number of Deaths")

```


## Which season of the year has the most gun deaths?

Assume that:

* Winter = January-March
* Spring = April-June
* Summer = July-September
* Fall = October-December

```{r season}
library(ggplot2) # load tidyverse packages, including ggplot2
library(readr)
library(knitr)
gun_deaths <- read_csv("gun_deaths.csv") 
data("gun_deaths") 

gd <- gun_deaths %>%
    select(year, race, education)
    group_by(education)
ggplot(gun_deaths, aes(x = sex, y = age)) +
 geom_boxplot(aes(fill = "black"))  +
 theme(legend.position =none) +
 labs(title="Number Gun Death Victims by Intent", 
           x="Intent",
       y="Number of Deaths")
```


# These are more open ended questions from here on out, you could look at making multiple summaries or types of plots. 

### Are whites who are killed by guns more likely to die because of suicide or homicide? How does this compare to blacks and hispanics?

```{r race}

```

(hint maybe looking at percentages would be good)
```{r race_2}

```


### Are police-involved gun deaths significantly different from other gun deaths? Assess the relationship between police involvement and age, police involvement and race, and the intersection of all three variables.

```{r police}

```

Think about performing stats such as a t.test to see if differences you are plotting are statistically significant

```{r stats?}


```


```{r police_2}

```

```{r stats_2?}


```


Finally, all together now:
```{r police_3}

```


## Session info

```{r}
# always good to have this for reproducibility purposes
devtools::session_info()
```

